---
title: "Simple document"
output: github_document
---

We can use gitignore to ignore the files that we don't want to commit, for example, write line "/data" so that we cannot commit the data folder. 

```{r, echo = FALSE, message = FALSE}
#we don't see anything here in the output, just things that we do ourselves in the background
library(tidyverse)
```

## Import some data

I want to import `FAS_litters.csv`;

```{r}
# taken directly from the data folder in our current directory
litters_df = read_csv("data/FAS_litters.csv")
```

```{r, include=FALSE}
#absolute path, should avoid
read_csv("/Users/Yma04/R/DS_P8105/Lecture Notes/p8105_data_wrangling_i/data/FAS_litters.csv")
#we can use "tab" after typing "/Users/", r will automatically go from there 
```

First step: clean the data to the format that I'm comfortable with
```{r}
names(litters_df)

litters_df = janitor::clean_names(litters_df)
```

first line of code show the original names of the df
second line use janitor to convert the names to snake cases, replacing the old version of the df with the new version, the one cleaned by janitor 
`janitor::clean_names(litters_df)` is a short way to tell r to find the clean_names function in janitor without having to load all of the janitor package 

## Taking a look at the data set

print out the dataset 
```{r}
litters_df 

head(litters_df)
tail(litters_df)
```

One I use a lot it: viewing the data set on a seperate sheet
```{r, eval=FALSE}
#viewing is interactive, so maybe it will have problem during knitting, so put eval=FALSE
view(litters_df)
```

Here's `skimr`:
```{r}
skimr::skim(litters_df)
```

## Arguments in `read_csv`

There are multiple arguments in `read_csv`, such as skip, na, etc. which can be helpful to clean the data, for example:

```{r, eval=FALSE}
#skip first five rolls, change colomn names(becomes x1,x2...), every time it sees a "Low8", regard it as missing data
litters_df =
  read_csv(
    "data/FAS_litters.csv",
    skip = 5,
    col_names = FALSE,
    na = "Low8"
  )
```

## Parsing columns

```{r}
# "./" means staring from here and go into this directory
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)
```

## Reading from Excel

```{r}
library(readxl) # usually put on top
```

reading MLB data.

```{r}
mlb11_df = read_excel("data/mlb11.xlsx")
```

LotR Words is next, the data set is not nicely organized, we need to import cells B3-D6

```{r}
fellow_df = read_excel("data/LotR_Words.xlsx",range = "B3:D6")
```

## Read a SAS file

```{r}
library(haven)
```

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```


