---
title: "Simple document"
output: github_document
---

We can use gitignore to ignore the files that we don't want to commit, for example, write line "/data" so that we cannot commit the data folder. 

```{r, echo = FALSE, message = FALSE}
#we don't see anything here in the output, just things that we do ourselves in the background
library(tidyverse)
```

## Import some data

I want to import `FAS_litters.csv`;

```{r}
# taken directly from the data folder in our current directory
litters_df = read_csv("data/FAS_litters.csv")
```

```{r, include=FALSE}
#absolute path, should avoid
read_csv("/Users/Yma04/R/DS_P8105/Lecture Notes/p8105_data_wrangling_i/data/FAS_litters.csv")
#we can use "tab" after typing "/Users/", r will automatically go from there 
```

First step: clean the data to the format that I'm comfortable with
```{r}
names(litters_df)

litters_df = janitor::clean_names(litters_df)
```

first line of code show the original names of the df
second line use janitor to convert the names to snake cases, replacing the old version of the df with the new version, the one cleaned by janitor 
`janitor::clean_names(litters_df)` is a short way to tell r to find the clean_names function in janitor without having to load all of the janitor package 

## Taking a look at the data set

print out the dataset 
```{r}
litters_df 

head(litters_df)
tail(litters_df)
```

One I use a lot it: viewing the data set on a seperate sheet
```{r, eval=FALSE}
#viewing is interactive, so maybe it will have problem during knitting, so put eval=FALSE
view(litters_df)
```

Here's `skimr`:
```{r}
skimr::skim(litters_df)
```



