---
title: "Tidy Data"
author: "Anna Ma"
date: "10/3/2021"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
#we don't see anything here in the output, just things that we do ourselves in the background
library(tidyverse)
library(readxl)
library(haven)
```

## Piviot longer

Load the PULSE data

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Let's try to pivot

```{r}
# if your variable share the same prefix and you don't need that prefix in everything, use names_prefix ="the part u don't want"
#
pulse_tidy = 
  pulse_df %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, #taking the columns from bl to 12m
    names_to = "visit",         #pivot the column to a variable names visit
    names_prefix = "bdi_score_", # getting rid of the prefix bdi_score_
    values_to = "bdi"   
    #pivot all the value from original column bl-12m to a variable names bdi
  ) %>% 
  mutate(
    visit = replace(visit, visit == "bl", "00m"), 
    #replacing bl with 00m to have consistent naming
    visit = factor(visit) # converting visit to a factor variable 
  )
view(pulse_tidy)
```

don't do `gather()`, always use `pivot_longer()`

## Pivot_wider example

make up a results data table

```{r}
analysis_df = 
  tibble(
    group = c("treatment", "treatment", "control","control"),
    time = c("a","b","a","b"),
    group_mean = c(4,8,3,6)
  )

analysis_df %>%
  pivot_wider(
    names_from = "time",
    values_from = "group_mean"
  ) %>% 
  knitr::kable() #formatting the table so it's easier read 
```

don't do `spread()`, always use `pivot_wider()`

## Biding rows

import the LotR movie words data sheet 

```{r}
fellowship_df = 
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_rings")

two_towers_df = 
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_df = 
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")

#instead of solving the untidiness in each one when we created them, we can combine all three together first and then tidy it 

lotr_df = 
  bind_rows(fellowship_df, two_towers_df, return_df) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words"
  ) %>%
  mutate(race = str_to_lower(race)) %>%  
#making all races to lowercase, if you want to keep the original and have a new variable with race in lowercase, write "mutate(race_lower = str_to_lower(race))" instead. 
  relocate(movie)
```

use `rbind()` or `bind_rows` ? 
pretty much the same thing, like `read.csv` vs `read_csv`. `rbind()` is worse, so don't use

## Joins

Look at FAS data. This imports and cleans litters and pups data

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  #tidy the data set first, separate the group and dose in the first variable 
  separate(group, into = c("dose", "day_of_tx"), 3) %>%   # 3 letters in and split
  relocate(litter_number) %>% 
  mutate(dose = str_to_lower(dose))

pups_df = 
  read_csv("data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  #replace the sex variable with f and m so it's more clear, note `1`, cannot just use 1 because r would think it's a integer, using `` will force r to recognize it as a variable instead of a number
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))
``` 

Let's joins these up!

```{r}
fas_df = 
  # join into the pups_df everything from the litters_df
  left_join(pups_df, litters_df, by = "litter_number") %>% 
  relocate(litter_number, dose, day_of_tx)
```

